---
title: "wk05_ggplot"
author: "Patricia Faundez"
date: "February 5, 2016"
output: html_document
---

```{r setwd students}
wd = 'env-info_hw'

# set working directory for Console (vs Rmd)
if (wd %in% list.files() & interactive()){
    setwd(wd)
}

# ensure working directory
if (basename(getwd()) != wd){
  stop(sprintf("WHOAH! Your working directory is not in '%s'!\n   getwd(): %s", wd, getwd()))
}
```

#Metadata thin.txt

"varname" "descrip"
"1" "startyear" "year of thin"
"2" "thin" "percent canopy removed"
"3" "postyears" "years since thinnning"
"4" "day" "current day"
"5" "month" "current month"
"6" "trans" "transpiration (mm/day)"
"7" "evap" "evaporation (mm/day)"
"8" "lai" "Leaf Area"
"9" "psn" "Photosynthesis (gC/m2/day)"
"10" "cpool" "Non Structural Carbohydrate (gC/m2)"
"11" "plantc" "Total Biomass (kgC/m2)"
"12" "year" "year"

```{r}
getwd()

library(readr)
library(readxl) # install.packages('readxl')
library(tidyr)
library(httr)   # install.packages('httr')
#library(plyr)   # install.packages('plyr')
library(dplyr)
library(ggplot2) 
library(knitr)       # functions for knitting Rmd documents to html
library(RColorBrewer)
library(gridExtra)


```

#Everything in http://docs.ggplot2.org/current/

#Questions in Chunks: 3,19,42

#my task I explored Aesthetics, Lines specified by slope and intercept, geom_abline: slope and intercept

#Comparing Biomass with other variables
```{r}
thin <- read_csv("~/Winter_2016/Environmental_Informatics/faundez.github.io/env-info_hw/thin.csv")
#scatter plot Biomass v/s Thinning
p1=ggplot(thin, aes(x=plantc, y=thin))+geom_point(col="brown", shape=1, size=rel(1))
#How does Biomass and Thinning is compare
p1=p1+labs(x="Biomass", y="Thinning") 
  p1 = p1+ ggtitle("How does Biomass and Thinning compare")
#coef(lm(thin ~ plantc, data = thin)) # Calculate slope and intercept of line of best fit
#p1 + geom_smooth(method = "lm", se = FALSE)                                  #NOT WORKING ????
p1 = p1 + geom_abline(intercept=73.59,slope=-11.08, colour="black", size=1)
p1
```


```{r}
#scatter plot Biomass v/s photosyntesis
p2=ggplot(thin, aes(x=plantc, y=psn))+geom_point(col="blue", shape=1, size=rel(1))
p2
#How does Biomass and Photosyntes is compare
p2=p2+labs(x="Biomass", y="Photosyntesis") 
  p2 = p2+ ggtitle("How does Biomass and Photosyntesis compare")
#coef(lm(psn ~ plantc, data = thin)) # Calculate slope and intercept of line of best fit
  p2 = p2 + geom_abline(intercept=-0.001,slope=0.0008, colour="black", size=1)
p2
```

```{r}
#how does Leaf Area and Biomass compare
#scatter plot Leaf Area v/s Biomass
p3=ggplot(thin, aes(x=plantc, y=lai))+geom_point(col="green", shape=1, size=rel(1))
p3=p3+labs(x="Biomass", y="Leaf Area") 
  p3 = p3+ ggtitle("How does Biomass and Leaf Area compare")
#coef(lm(lai ~ plantc, data = thin)) # Calculate slope and intercept of line of best fit
p3 = p3 + geom_abline(intercept=-1.5,slope=1.34, colour="black", size=1)
p3
 


```

```{r}
#scatter plot Biomass v/s Transpiration
p4=ggplot(thin, aes(x=plantc, y=trans))+geom_point(col="orange", shape=1, size=rel(1))
#How does Biomass and Transpiration is compare
p4=p4+labs(x="Biomass", y="Transpiration") 
  p4 = p4+ ggtitle("How does Biomass and Transpiration compare")
#coef(lm(trans ~ plantc, data = thin)) # Calculate slope and intercept of line of best fit
  p4 = p4 + geom_abline(intercept=-1.337,slope=0.67, colour="black", size=1)
p4
```

```{r}
#scatter plot Biomass v/s Evaporation
p5=ggplot(thin, aes(x=plantc, y=evap))+geom_point(col="orange", shape=1, size=rel(1))
#How does Biomass and Evaporation is compare
p5=p5+labs(x="Biomass", y="Evaporation") 
  p5 = p5+ ggtitle("How does Biomass and Evaporation compare")
#coef(lm(evap ~ plantc, data = thin)) # Calculate slope and intercept of line of best fit
  p5 = p5 + geom_abline(intercept=0.14,slope=-0.019, colour="black", size=1)
p5
```

```{r}
grid.arrange(p1,p2,p3,p4,p5)
```

#Comparing Biomass with other variables
```{r}
#scatter plot Non Structural Carbohydrate v/s Thinning
a1=ggplot(thin, aes(x=cpool, y=thin))+geom_point(col="brown", shape=1, size=rel(1))
#How does Non Structural Carbohydrate and Thinning is compare
a1=a1+labs(x="Non Structural Carbohydrate", y="Thinning") 
  a1 = a1+ ggtitle("How does Non Structural Carbohydrate and Thinning compare")
#coef(lm(thin ~ cpool, data = thin)) # Calculate slope and intercept of line of best fit
  a1 = a1 + geom_abline(intercept=39.7,slope=-25.4, colour="black", size=1)
a1
```


```{r}
#scatter plot Non Structural Carbohydrate v/s photosyntesis
a2=ggplot(thin, aes(x=cpool, y=psn))+geom_point(col="blue", shape=1, size=rel(1))
#How does Non Structural Carbohydrate and Photosyntes is compare
a2=a2+labs(x="Non Structural Carbohydrate", y="Photosyntesis") 
  a2 = a2+ ggtitle("How does Non Structural Carbohydrate and Photosyntesis compare")
#coef(lm(psn ~ cpool, data = thin)) # Calculate slope and intercept of line of best fit
  a2 = a2 + geom_abline(intercept=0.0006,slope=0.0034, colour="black", size=1)
a2
```

```{r}
#how does Leaf Area and Non Structural Carbohydrate compare
#scatter plot Leaf Area v/s Non Structural Carbohydrate
a3=ggplot(thin, aes(x=cpool, y=lai))+geom_point(col="green", shape=1, size=rel(1))
a3=a3+labs(x="Non Structural Carbohydrate", y="Leaf Area") 
  a3 = a3+ ggtitle("How does Non Structural Carbohydrate and Leaf Area compare")
#coef(lm(lai ~ cpool, data = thin)) # Calculate slope and intercept of line of best fit
a3 = a3 + geom_abline(intercept=3.49,slope=1.46, colour="black", size=1)
a3

```

```{r}
#scatter plot Non Structural Carbohydrate v/s Transpiration
a4=ggplot(thin, aes(x=cpool, y=trans))+geom_point(col="orange", shape=1, size=rel(1))
#How does Non Structural Carbohydrate and Transpiration is compare
a4=a4+labs(x="Non Structural Carbohydrate", y="Transpiration") 
  a4 = a4+ ggtitle("How does Non Structural Carbohydrate and Transpiration compare")
#coef(lm(trans ~ cpool, data = thin)) # Calculate slope and intercept of line of best fit
  a4 = a4 + geom_abline(intercept=1.109,slope=0.914, colour="black", size=1)
a4
```

```{r}
#scatter plot Non Structural Carbohydrate v/s Evaporation
a5=ggplot(thin, aes(x=cpool, y=evap))+geom_point(col="orange", shape=1, size=rel(1))
#How does Non Structural Carbohydrate and Evaporation is compare
a5=a5+labs(x="Non Structural Carbohydrate", y="Evaporation") 
  a5 = a5+ ggtitle("How does Non Structural Carbohydrate and Evaporation compare")
#coef(lm(evap ~ cpool, data = thin)) # Calculate slope and intercept of line of best fit
  a5 = a5 + geom_abline(intercept=0.016,slope=0.08, colour="black", size=1)
a5
```

```{r}
grid.arrange(a1,a2,a3,a4,a5)
```

#Aesthetics, Lines: horizontal and vertical geom_vline: xintercept, geom_hline: yintercept


```{r}

p <- ggplot(thin, aes(x=as.factor(postyears), y=plantc)) + geom_point()
p
# Fixed values
p + geom_vline(xintercept = 10)
```

```{r}
p + geom_hline(yintercept = 2:6)
```

```{r}
p + geom_hline(yintercept = 4)
```

```{r}
p + geom_abline() # Can't see it - outside the range of the data
p + geom_abline(intercept = 4)
```


#Copy examples from the class with thin data

```{r}
# basic plot
z = ggplot(thin, aes(y=plantc, x=month))+geom_point()
z
z = z + labs(y="Biomass (kgC/m2)", x="Month")
zclr = z+theme(axis.text= element_text(face="bold", colour="blue", size=12))
z
zclr
grid.arrange(z,zclr)
```

```{r}
# note the use of as factor (boxplot need factors to organize)
z1=ggplot(thin, aes(y=plantc, x=as.factor(month)))+geom_boxplot(colour="blue", fill="yellow")
z1=z1+labs(y="Biomass (kgC/m2)", x="Month")
z1

```


```{r}
# basic plot
z = ggplot(thin, aes(y=evap, x=month))+geom_point()
z
z = z + labs(y="Evaporation", x="Month")
zclr = z+theme(axis.text= element_text(face="bold", colour="blue", size=14))
z
zclr
grid.arrange(z,zclr)
```


```{r}
z2 = ggplot(thin, aes(x=year,y=evap))+stat_summary(fun.y="sum", geom="bar", col="blue")
z2
```

```{r}
z2 = ggplot(thin, aes(x=as.factor(month),y=evap))+stat_summary(fun.y="mean", geom="bar")

z2 = z2 + scale_x_discrete(labels=abbreviate) + labs(x="Months",y="Evaporation")
z2
```

```{r}
z2 = z2 + facet_wrap(~year)
z2
```

```{r}
z3 = z2 + theme(axis.text=element_blank(), strip.text = element_blank(), strip.background = element_blank())
z3
```


```{r, echo=FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(readr)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
})
```


#ggplot(data=co2_percapita, aes(y=n, x=Country))+stat_summary(fun.y="mean", geom="point") + labs(x="Time since thinning", y="Carbon Sequestration")

### per capita CO<sub>2</sub> emissions 

You'll join the [population](http://data.okfn.org/data/core/population) dataset to calculate per capita CO<sub>2</sub> emissions.

_**Task**. Calculate the per capita emissions by country (not World or EU28) and return the top 5 emitting countries for **2014**._

_**Task**. Summarize the per capita emissions by country (not World or EU28) as the **mean** (ie average)  value across all years and return the top 5 emitting countries._ 


```{r read co2 & popn, eval=T}
suppressPackageStartupMessages({
  library(readr)
  library(readxl) # install.packages('readxl')
  library(dplyr)
  library(tidyr)
  library(httr)   # install.packages('httr')
  library(ggplot2)
  library(knitr)
})

# co2 Excel file
url = 'http://edgar.jrc.ec.europa.eu/news_docs/CO2_1970-2014_dataset_of_CO2_report_2015.xls'
xls = 'co2_europa.xls'
if (!file.exists(xls)) writeBin(content(GET(url), 'raw'), xls)

# read in carbon dioxide emissions file
co2 = read_excel(xls, skip=12) %>%
  gather(year, ktons_co2, -Country) %>%
  filter(!Country %in% c('World','EU28') & !is.na(Country)) %>%
  mutate(year = as.integer(as.character(year))) %>%
  rename(country=Country)
co2

# read in population
popn = read_csv('https://raw.githubusercontent.com/datasets/population/master/data/population.csv') %>%
  select(country=`Country Name`, year=Year, popn=Value) 
popn
```

In order to calculate per capita emissions, you'll need to join the `popn` data frame with the `co2` data frame before calculating something like `ktons_co2 / popn`.

One of the trickiest parts about joining data is ensuring that values in the key columns are matching. The column `year` is a simple 4 digit integer in both tables (after converting to `as.integer()` for the gathered `co2`). But what about `country`?

Let's look at just `year==2014` to see what values of `country` in table `co2` are not found in table `popn`, ie `anti_join()`:

```{r co2$popn not in popn$country, eval=T}
# co2$popn not in popn$country
co2 %>%
  filter(year==2014) %>%
  anti_join(
    popn %>%
      filter(year==2014),
    by='country') %>%
  arrange(desc(ktons_co2))
```

Whoah, we're missing 47 "countries", including the "United States of America"! So what is "United States of America" called in  `popn$country`?

```{r popn$country not in co2$popn, eval=T}
# popn$country not in co2$popn
popn %>%
  filter(year==2014) %>%
  anti_join(
    co2 %>%
      filter(year==2014), 
    by='country') %>%
  arrange(desc(country))
```

Ok, so in `popn$country` we have "United States" which doesn't get matched with "United States of America" in `co2$country`? How to fix this? We need a lookup table for translating country values between tables. I went ahead and made one for you by copying the two anti_join tables above and matching rows where possible.

```{r cntry, eval=T}
# get lookup table to translate popn$country to co2$country
cntry = read_csv('https://raw.githubusercontent.com/ucsb-bren/env-info/gh-pages/data/co2_country_to_popn.csv')
cntry
```

So before we join the tables `co2` and `popn`, first translate country in one to match the other.

```{r update co2$country to popn$country, eval=T}
# update co2$country to popn$country
co2_m = co2 %>%
  left_join(
    cntry, 
    # note use of named vector to columns with different names
    by = c('country'='country_co2')) %>%    
  mutate(
    # note use of ifelse to upated country with popn$country_popn
    country = ifelse(!is.na(country_popn), country_popn, country))

# check that "United States of America" changed to "United States"
co2_m %>%
  filter(year==2014) %>%  #why this is not working???? I did it again in next chuck  ???????
  arrange(desc(country)) %>%
  head(15)
```

Ok, so `co2_m$country` got updated with values to match `popn$country`, eg "United States of America" to "United States".

Please use this matching `co2_m` to join with `popn` and calculate per capita emissions.


```{r}

co2_percapita = co2_m %>%
  filter(year > 2013) %>%
  filter(country != c("World", "EU28")) %>%
  left_join(
    popn, by = c("country", "year")) %>%
  mutate(
    ratio = ktons_co2 / popn) %>% 
  select(country, ratio)

#h1 = ggplot(data=co2_percapita, aes(y=ratio, x=country))+geom_point()+ labs(x="Country", y="ratio (CO2(metric tons) / population)")+ ggtitle("Emissions of CO2 per capita per country")
#h1
#h1 + ylim(0.02, 0.18)

co2_percapita_0.02 = co2_percapita %>%
  filter(ratio > 0.015)
t = ggplot(data=co2_percapita_0.02, aes(y=ratio, x=country))+geom_point()+ labs(x="Country", y="ratio (CO2(metric tons) / population)")+ ggtitle("Emissions of CO2 per capita per country")
t = t+theme_bw()
t
#how to fix country names vertical in axis x??????

```

