---
title: "wk05_ggplot"
author: "Patricia Faundez"
date: "February 5, 2016"
output: html_document
---


```{r setup, echo=F}
# NOTE: This chunk sets the default to eval=F for all following chunks.
#       Am just using this for instructions. Please exclude in your assignment to see all results.
#knitr::opts_chunk$set(eval=FALSE)
```


### RStudio Shortcuts

- [rstudio-IDE-cheatsheet.pdf](./refs/cheatsheets/rstudio-IDE-cheatsheet.pdf)
    - Show Keyboard Shortcuts `?`: Alt+Shift+**K** (Win), Option+Shift+**K** (Mac)
    - (Un)Comment lines `# `: Ctrl+Shift+**C** (Win), Cmd+Shift+**C** (Mac)
    - Insert `%>%`: Ctrl+Shift+**M** (Win), Cmd+Shift+**M** (Mac)
    - Attempt completion: **Tab** or Ctrl+Space (Win), **Tab** or Cmd+Space (Mac), 
    - Knit Document: Ctrl+Shift+**K** (Win), Cmd+Shift+**K** (Mac)
    - Copy Lines Up/Down: Shift+Alt+**$\uparrow$**/**$\downarrow$** (Win), Cmd+Option+**$\uparrow$**/**$\downarrow$** (Mac)


### Individual Assignment `wk05_ggplot`

## joining data

### setup

Ensure that you're in the same working directory `env-info_hw` when you Knit HTML as when you test code in the Console.

```{r setwd students}
wd = 'env-info_hw'

# set working directory for Console (vs Rmd)
if (wd %in% list.files() & interactive()){
    setwd(wd)
}

# ensure working directory
if (basename(getwd()) != wd){
  stop(sprintf("WHOAH! Your working directory is not in '%s'!\n   getwd(): %s", wd, getwd()))
}
```

### `bind_cols`

```{r bind_cols}
library(dplyr)
y = data.frame(
  x1 = c('A','B','C'), 
  x2 = c( 1 , 2 , 3), 
  stringsAsFactors=F)
z = data.frame(
  x1 = c('B','C','D'), 
  x2 = c( 2 , 3 , 4), 
  stringsAsFactors=F)
y
z
bind_cols(y, z)
```

### `bind_rows`

```{r bind_rows}
y
z
bind_rows(y, z)
```

### `union`

```{r union}
y
z
union(y, z)
```

### `intersect`

```{r intersect}
y
z
intersect(y, z) #only things in commom
```

### `setdiff`

```{r setdiff}
y
z
setdiff(y, z)
```

### `left_join`

```{r left_join}
songs = data.frame(
  song = c('Across the Universe','Come Together', 'Hello, Goodbye', 'Peggy Sue'),
  name = c('John','John','Paul','Buddy'), 
  stringsAsFactors=F)
artists = data.frame(
  name = c('George','John','Paul','Ringo'),
  plays = c('sitar','guitar','bass','drums'), 
  stringsAsFactors=F)
left_join(songs, artists, by='name')
```

### `inner_join`

```{r inner_join}
inner_join(songs, artists, by = "name")
```

### `semi_join`

```{r semi_join}
semi_join(songs, artists, by = "name")
```

### `anti_join`

```{r anti_join}
anti_join(songs, artists, by = "name")
```

### per capita CO<sub>2</sub> emissions 

You'll join the [population](http://data.okfn.org/data/core/population) dataset to calculate per capita CO<sub>2</sub> emissions.

_**Task**. Calculate the per capita emissions by country (not World or EU28) and return the top 5 emitting countries for **2014**._

```{r read co2 & popn}
library(readr)
library(readxl) # install.packages('readxl')
library(dplyr)
library(tidyr)
library(httr)   # install.packages('httr')
library(plyr)   # install.packages('plyr')

print(getwd())

# population Excel file
url = 'http://edgar.jrc.ec.europa.eu/news_docs/CO2_1970-2014_dataset_of_CO2_report_2015.xls'
out = 'co2_europa.xls'
if (!file.exists(out)) writeBin(content(GET(url), 'raw'), out)

# country lookup csv file
url = 'https://raw.githubusercontent.com/ucsb-bren/env-info/gh-pages/data/co2_country_to_popn.csv'
out = 'co2_country_to_popn.csv'
if (!file.exists(out)) writeBin(content(GET(url), 'raw'), out)

# read in carbon dioxide emissions file
xls = 'co2_europa.xls'
co2 = read_excel(xls, skip=12)
# to match names use anti-join  

# get lookup table to translate some country names in popn to those in co2 table
cntry = read_csv('co2_country_to_popn.csv')

# read in population and revalue some country names to match co2 table
popn = read_csv('https://raw.githubusercontent.com/datasets/population/master/data/population.csv') %>%
  select(country=`Country Name`, year=Year, popn=Value) %>%
  mutate(
    country = plyr::revalue( #revalue country with any matches between countries
      country, 
      setNames(cntry$country_co2, cntry$country_popn)))
```

###Getting emission's countries in 2014

```{r gather}
co2
d = gather(co2, "year", "n", -1)
d
# Other way to do the same: gather(co2, "year", "n", 2:ncol(co2)) or gather(co2, "year", "n", -Country)

#Converting my year column from factor to numeric

e = mutate(d, year = as.numeric(as.character(year)))
e

new2 = filter(e, year > 2013)   # other way is new <- e[e$year == 2014,]

new3 = arrange(new2, desc(n))

co2_2014 = filter(new3, Country != c("World", "EU28"))

head(co2_2014)

```

###Getting population's countries in 2014

```{r}
pop_2014 = filter(popn, year > 2013) 

names(pop_2014)[1]<-paste("Country") #changing the name in the first column from "country" to "Country", that it will match with the co2_2014 for joinning purposes. 
```

```{r}
Per_Capita = left_join(co2_2014, pop_2014, by = "Country")

co2_percapita = mutate(Per_Capita, ratio = n / popn)

arrange(co2_percapita, desc(ratio))

head(co2_percapita)

```

```{r, echo=FALSE, warning=FALSE}

suppressWarnings(suppressMessages(library(readr)))
suppressWarnings(suppressMessages(library(dplyr)))
suppressWarnings(suppressMessages(library(tidyr)))
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(knitr)))

```

```{r}
ggplot(data=co2_percapita, aes(y=ratio, x=Country))+geom_point()
```

```{r}
ggplot(data=co2_percapita, aes(y=n, x=Country))+geom_point()
```

_**Task**. Summarize the per capita emissions by country (not World or EU28) as the mean (ie average)  value across all years and return the top 5 emitting countries._ 

```{r}
e
ee = filter(e, Country != c("World", "EU28"))

names(popn)[1]<-paste("Country")
popn

allyears = left_join(ee, popn, by = c("Country", "year"))
allyears

pc_allyears = mutate(allyears, ratio = n / popn)
pc_allyears

n = pc_allyears %>% select(Country, ratio)

n <- n[!is.na(n$ratio), ]

n <- aggregate(n$ratio, by = list(n$Country), FUN = mean)
names(n) <- c("Country", "mean")
m = arrange(n, desc(mean))
head(m)
   
```

#I tried dplyr but make me this error

> n %>% group_by(Country)%>%
+ summarise(
+ mean = mean(ratio))%>%
+ write_csv("x")
Error in mean(ratio) : object 'ratio' not found
> head(x)
Error in head(x) : object 'x' not found

n %>% group_by(Country)%>% 
  summarise(
    mean = mean(ratio))%>% 
    write_csv("x")
    head(x)
 
