---
title: "wk06_widgets"
author: "Patricia Faundez"
date: "February 12, 2016"
output: html_document
---

### setup

Ensure that you're in the same working directory `env-info_hw` when you Knit HTML as when you test code in the Console.

```{r setwd env-info_hw, eval=FALSE}
wd = 'env-info_hw'

# set working directory for Console (vs Rmd)
if (wd %in% list.files() & interactive()){
    setwd(wd)
}

# ensure working directory
if (basename(getwd()) != wd){
  stop(sprintf("WHOAH! Your working directory is not in '%s'!\n   getwd(): %s", wd, getwd()))

}

# set default eval option for R code chunks
#knitr::opts_chunk$set(eval=FALSE)
```


### `manipulate`


**Task.** Add another R chunk with a slider adjusting the number of bins from 5 to 50, with step increments of 5.


```{r manipulate, eval=F}
library(manipulate) # install.packages('manipulate')

manipulate({
  
  faithful %>%
    ggplot(aes(eruptions)) + 
    geom_histogram(aes(y = ..density..),  bins = 20) +
    geom_density(color='blue', size=2, adjust=a) +
    xlab('duration (minutes)') +
    ylab('frequency density') +
    ggtitle('geyser eruption duration')
  
}, a = slider(min = 5, max = 50, initial = 5, label = 'bandwidth adjustment', step = 5))
```



### `ggvis`


**Task.** Add another R chunk that only applies the `add_tooltip` on mouse click.

```{r}
library(dplyr)
library(ggvis)
cars = mtcars %>%
  add_rownames('model') %>%        # dplyr drops rownames
  mutate(id = row_number())       # add an id column to use ask the key

all_values <- function(x) {
  if(is.null(x)) return(NULL)
  row <- cars[cars$id == x$id, ]
  paste0(names(row), ": ", format(row), collapse = "<br/>")
}

cars %>% 
  ggvis(x = ~wt, y = ~mpg, key := ~id) %>%
  layer_points() %>%
  add_tooltip(all_values, 'click')
```


## htmlwidgets


**Task.** Output the table again with `datatable` and set the `options` to have `pagelength` of just 5 rows. (See `?datatable` and <http://rstudio.github.io/DT/>).

```{r}
library(dplyr)
library(knitr) # install.packages('kable')
library(DT) # install.packages('DT')
datatable(iris)

# remove document elements
datatable(iris, options = list(pageLength = 5))
```


